(a) What is the difference between categories 5 and 14?

5 End of line (<return>)
14 Comment character (%)
Everything between a comment character and the beginning of the next line is dropped.
The end of line is either dropped (if preceeded by spaces) or changed into a space.

Answer:

Both characters terminate the current line; but a character of category 5
might be converted into 10 or a par token, while a character of category 14 never
produces a token.


(b) What is the difference between categories 3 and 4? 

3 is to shifted from/to math mode.
4 is to align things in an array.

Answer:

They produce character tokens stamped with different category
numbers. For example, $ 3 is not the same token as $ 4 , so TEX’s digestive processes
will treat them differently.


(c) What is the difference between categories 11 and 12?

Letters are letters, other characters are... others.

Answer:

Same as (b), plus the fact that control sequence names
treat letters differently.


(d) Are spaces ignored after active characters?

No.


(e) When a line ends with a
comment character like %, are spaces ignored at the beginning of the next line?

No, because the end of line has been dropped, and not transformed into space.
The first space is passed, subsequent ones are ignored.

Answer:

Yes; characters of category 10 are ignored at the
beginning of every line, since every line starts in state N .


(f) Can
an ignored character appear in the midst of a control sequence name?

The control word stops before the first non-letter.

Answer:

No.

\bye


